---
title: "Case Study 2"
author: "Chad Kwong"
date: "11/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn = -1)
library(ggplot2)
library(tidyr)
library(dplyr)
library(e1071)
library(class)
library(blorr)
library(caret)
library(lattice)
library(stats)
library(plotly)
library(groupdata2)
```

# YouTube Link: 
https://www.youtube.com/watch?v=Iuw4K7tsoXE

# Executive Summary:
The following markdown file follows the exploratory data analysis performed on the employee information data set provided, titled CaseStudy2-data.csv. The purpose of this EDA is to predict Attrition and Monthly income based on the variables available within the data set. Any questions can be addressed to Chad Kwong at ckwong@smu.edu.

# Introduction: 
The markdown file starts with the initial loading/cleaning of the data and moves on to each aspect of the EDA. Each chunk of code is commented and a summary outline is provided before and after each chunk as well. Any questions can be addressed to Chad Kwong at ckwong@smu.edu.

# Loading and cleaning up the data
```{r}
# Initial Loading of data
AttritionData = read.csv("/Users/chadkwong/Desktop/SMU/GitHub/Case-Study-2/CaseStudy2-data.csv")
str(AttritionData)

# Cleaning and Altering data for initial investigation
AttritionData$Sorted_MonthlyIncome = cut(AttritionData$MonthlyIncome, breaks = c(0,4000,8000,12000,16000,20000), labels = c("Under 4k", "4k-8k", "8k-12k", "12k-16k", "16k+"))
AttritionData$Sorted_YearsAtCompany = cut(AttritionData$YearsAtCompany, breaks = c(-1,4,8,12,16,max(AttritionData$YearsAtCompany)), labels = c("Under 4","4-8","8-12","12-16","16+"))
AttritionData$Attrition = factor(AttritionData$Attrition)
AttritionData$JobRole = factor(AttritionData$JobRole)
AttritionData$TotalWorkingYears = factor(AttritionData$TotalWorkingYears)
AttritionData$YearsAtCompany = factor(AttritionData$YearsAtCompany)
AttritionData$YearsWithCurrManager = factor(AttritionData$YearsWithCurrManager)
yes = which(AttritionData$Attrition == "Yes")
AttritionYes = AttritionData[yes,]
table(is.na(AttritionData))
```
No missing values were found in the data set.


# Graphs included in initial investigation
The following code outlines my thought process around my initial investigation into predicting attrition based on 3 variables. This investigation was performed using empirical interpretation without supporting evidence.
```{r}
# Box plot of Stock Option Level and Monthly Income based on all Yes responses to Attrition
StockIncomeBox = ggplot(data = AttritionYes, aes(x = StockOptionLevel, y = MonthlyIncome)) + geom_boxplot()

# plots exploring relationship between Stock Option Level and Monthly Income
StockHist = ggplot(AttritionData, aes(x = StockOptionLevel, fill = Sorted_MonthlyIncome)) + geom_histogram(stat = "count") + facet_wrap(~Attrition)
StockHist
StockIncomePoint = ggplot(data = AttritionData, aes(x = StockOptionLevel, y = MonthlyIncome, color = Sorted_MonthlyIncome)) + geom_point() + facet_wrap(~Attrition)

# exploring distribution of monthly income and relationship to job level
IncomeHist = ggplot(AttritionData, aes(x = Sorted_MonthlyIncome, fill = Sorted_MonthlyIncome)) + geom_histogram(stat = "count") + facet_wrap(~Attrition)
IncomeHist
SortedIncomePoint = ggplot(AttritionData, aes(x = JobLevel, y = MonthlyIncome, color = Sorted_MonthlyIncome)) + geom_point()
SortedIncomeBox = ggplot(AttritionData, aes(x = MonthlyIncome, y = JobLevel, color = Sorted_MonthlyIncome)) + geom_boxplot()

# Histogram of years at company divided by Attrition
YearHist = ggplot(AttritionData, aes(x = Sorted_YearsAtCompany, fill = Sorted_YearsAtCompany)) + geom_histogram(stat = "count") + facet_wrap(~Attrition)
YearHist

# Histogram of Education divided by Attrition
EducationHist = ggplot(AttritionData, aes(x = AttritionData$Education)) + geom_histogram(stat = "count") + facet_wrap(~Attrition)

# Histogram of Job Satisfaction divided by Attrition
JobSatisfHist = ggplot(AttritionData, aes(x = AttritionData$JobSatisfaction)) + geom_histogram(stat = "count") + facet_wrap(~Attrition)

# Histogram of Job Roles
JobRoleHist = ggplot(AttritionData, aes(x = AttritionData$JobRole)) + geom_histogram(stat = "count") + facet_wrap(~Attrition)

# Histogram of Environment Satisfaction
environmentsat = ggplot(AttritionData, aes(x = AttritionData$EnvironmentSatisfaction)) + geom_histogram(stat = "count") + facet_wrap(~Attrition)


```
With all the plots above in mind, I chose to further investigate Monthly Income, Stock Option level, and Years at Company.


# Verifying investigation
I chose to attempt to use Learning Vector Quantization algorithm to compute the importance of variables within the attrition data set and use a repeated cross validation for verifying the model.
```{r}
# Using LVQ to predict variables and then performing a cross validation. Importance of the variables are then computed and displayed on a graph

#creating and training the model
  model <- train(Attrition~.,  data=AttritionData[,c(2,3,5,7,8,11,12,14,15,16,18,20,21,22,25,26,27,29,30,31,32,33,34,35,36)], method="lvq", preProcess="scale", trControl=trainControl(method="cv"))

# estimating variable importance using the model
  importance <- varImp(model, scale=FALSE)
  
# summarizing the importance
  print(importance)
  
# plotting the importance
  plot(importance, main = "Ranking of Variables")
```

The graph above predicts that the top 3 contributing variables are Monthly Income, Total Working Years, and Years at Company when predicting Attrition.


# Creating a Naive Bayes Model
I chose to create a Naive Bayes model using the Monthly Income, Total Working Years, and Years at company. I then wanted to create a model using the best possible seed, so I calculated the best seed based off of the first 50 seeds. I then choose to compare the model using the predicted variables to the model created using my own chosen variables and then build the superior model with the optimum seed value.
```{r}
## Variables to use: Monthly Income(c20)
## Considerable Variables: Stock Option Level(c29) Years at Company(c33), Total Working Years(c30)

# Testing Accuracy of top 3 predicted variables: Monthly Income, Total Working Years, Years at Company
nums = data.frame(accuracy = numeric(50), seed = numeric(50))
for(i in 1:50)
{
  set.seed(i)
  Indices = sample(seq(1:length(AttritionData$Attrition)),round(.70*length(AttritionData$Attrition)))
  train = AttritionData[Indices,]
  test = AttritionData[-Indices,]
  train = upsample(train, cat_col = "Attrition")
  train = upsample(train, cat_col = "Attrition")
  model = naiveBayes(Attrition ~ MonthlyIncome + TotalWorkingYears + YearsAtCompany, data = train[,c(3,20,33,30)], laplace = 10)
  predictions = predict(model, test)
  CM = confusionMatrix(predictions,test$Attrition)
  nums$accuracy[i] = CM$overall[1]
  nums$seed[i] = i
  nums$Sensitivity[i] = CM$byClass[1]
  nums$Specificity[i] = CM$byClass[2]
}
# Displaying row with highest accuracy
nums[which(nums$accuracy == max(nums$accuracy)),]
# Displaying Confusion Matrix Statistics for all seeds 1-50
legendcolors <- c("Accuracy" = "#AB0E14", "Sensitivity" = "#F1B11B", "Specificity" = "#957531")
ggplot(nums) + geom_line(aes(x = seed, y = accuracy, color = "Accuracy")) + geom_line(aes(x = seed, y = Sensitivity, color = "Sensitivity")) + geom_line(aes(x = seed, y = Specificity, color = "Specificity")) + labs(x = "Seed", y ="Score", color = "Legend") + scale_color_manual(values = legendcolors) + ggtitle("Confusion Matrix Statistics for 50 Seeds")

# Testing Accuracy of variables chosen from initial investigation: Monthly Income, Stock Option Level, Years at Company
accs = data.frame(accuracy = numeric(50), seed = numeric(50))
for(i in 1:50)
{
  set.seed(i)
  Indices = sample(seq(1:length(AttritionData$Attrition)),round(.70*length(AttritionData$Attrition)))
  train = AttritionData[Indices,]
  test = AttritionData[-Indices,]
  train = upsample(train, cat_col = "Attrition")
  model = naiveBayes(Attrition ~ ., data = train[,c(3,20,33,29)])
  predictions = predict(model, test)
  CM = confusionMatrix(predictions,test$Attrition)
  accs$accuracy[i] = CM$overall[1]
  accs$seed[i] = i
  accs$Sensitivity[i] = CM$byClass[1]
  accs$Specificity[i] = CM$byClass[2]
}
# Displaying highest accuracy for comparison
accs[which(accs$accuracy == max(accs$accuracy)),]

## Model Using Monthly Income, Total Working Years, and Years at Company with seed 20
set.seed(20)
#creating training and test sets
Indices = sample(seq(1:length(AttritionData$Attrition)),round(.70*length(AttritionData$Attrition)))
train = AttritionData[Indices,]
test = AttritionData[-Indices,]
#upsampling Attrition column for model
train = upsample(train, cat_col = "Attrition")
train = upsample(train, cat_col = "Attrition")
model = naiveBayes(Attrition ~ ., data = train[,c(3,20,33,30)])
predictions = predict(model, test)
confusionMatrix(predictions,test$Attrition)
# Writing attrition predictions to a new CSV file with the test data set
test$PredictedAttrition = predictions
AttritionPredictions = test[,c(1,3,20,30,33,39)]
write.csv(AttritionPredictions,"/Users/chadkwong/Desktop/SMU/GitHub/Case-Study-2/Attrition Predictions.csv", row.names = FALSE)
```
The above optimized Naive Bayes model provides an accuracy of ~75%, sensitivity of ~.7778, and specificity of ~.6111. With a small P-value of 1.2e-05, there is enough evidence to support the claim that Monthly Income, Years at Company, and Total Working Years are considerable variables when predicting Attrition.


# Predicting Monthly Income Using Variables from Model
When predicting monthly income, I chose to use the variables used in my previous model NB model. Using a linear regression model, I predicted the relationship between Monthly Income, Years at Company, and Total Working Years. I then perform a Leave One Out Cross Validation on the fit used and display the relationship between all 3 variables within a 3d scatter plot.
```{r, warning = FALSE}
# Regression Model for Predicting monthly income
fit = lm(MonthlyIncome ~ TotalWorkingYears + YearsAtCompany, data = AttritionData)
summary(fit)
preds = predict(fit)

# Performing LOO Cross Validation for both models
train(MonthlyIncome ~ TotalWorkingYears + YearsAtCompany, data = AttritionData, method = "lm", trControl = trainControl(method = "LOOCV"))

surface_plot <- plot_ly(AttritionData, 
                     x = ~TotalWorkingYears,
                     y = ~YearsAtCompany, 
                     z = ~MonthlyIncome,
                     name = "Attrition Data", 
                     type = "scatter3d",
                     mode = "markers",
                     size = .75)

# Adding Title to Graph and proper labels for axis
surface_plot <- surface_plot %>% layout(title = "Relationship Between Monthly Income, Years At Company, and Total Working Years", xaxis = list(title = 'Total Working Years',zeroline = TRUE), yaxis = list(title = 'Years at  Company'), zaxis = list(title = 'Monthly Income'))

# Adding in predicted Monthly Income
surface_plot <-surface_plot %>% add_trace(z = ~preds, name = "Predicted Monthly Income", mode = "markers")

# Displaying Plot
surface_plot
```
Using the model, we can observe a RMSE of ~2743 which is within the threshold of 3000. A p-value of 2.2e-16 is observed as well. From the plot above, we can also see that the predicted monthly income falls within the distribution of the actual data set. I can thus conclude that the fit supports the relationship between Monthly Income, Years at Company, and Total Working Years. However, I speculate that there is not enough information to confirm that Monthly income is solely dependent on Years At Company and Total Years Working